\pkg{cubble}: An R Package for Structuring Spatio-temporal Data

### 2022-02-08 2-the-cubble-pkg.Rmd
add transformation of calendar plot and projection

### sf

\textcolor{red}{add content}

>>>>

### sf and s2

Cubble is compatible with simple feature geometry list-column (\code{sfc}) in \pkg{sf} and  \code{s2_lnglat} vector in \pkg{S2} in the nested form. After creating a cubble, analysts can add the geometry vectors as needed to perform spatial operations:

```{r echo = TRUE}
library(s2)
# create a cubble
cb <- climate_flat %>% 
  cubble::as_cubble(key = id, index = date, coords = c(long, lat))

# western australia map
wa <- ozmaps::abs_ste %>% filter(NAME == "Western Australia")

# create a `s2_lnglat` vector based on long and lat and filter stations in wa
cb %>%
  mutate(ll = s2_lnglat(long, lat)) %>% 
  filter(s2_within(ll, wa))
```

```{r echo = TRUE}
library(sf)
# simplify australia map
aus <- ozmaps::abs_ste %>% rmapshaper::ms_simplify(keep = 0.01)

dt <- cb %>%
  mutate(
    # create `sfc` column based on long and lat
    ll = st_sfc(
      purrr::map2(long, lat, ~st_point(c(.x, .y))),
      crs = st_crs(aus)),
    # append state multi-polygon based on the `sfc` created
    state = aus$geometry[st_within(ll, aus, sparse = FALSE)],
    # adopt a different projection: lambert conformal conic (EPSG:3112)
    state = st_transform(state, crs = "EPSG:3112"))

dt
```

### 2022-02-08 3-advanced.Rmd
some changes as per meeting today
add transformation of calendar plot and projection

# Advanced features/ considerations

@@

## Glyph map

Glyph map [@Wickham2012-yr] plots the time series as single glyph on the map and in \proglang{R}, \pkg{GGally} implements the glyph map through the \code{glyphs()} function. The function constructs a data frame with calculated position (\code{gx}. \code{gy}. \code{gid}) of each point on the time series using linear algebra (Equation 1 and 2 in @Wickham2012-yr). The data can then be piped into \code{ggplot} to create the glyph map as:

>>>>

# Other features and considerations

@@

## Spatio-temporal transformations

Spatio-temporal transformation is a useful technique to extract information form spatio-temporal data. Glyph maps [@Wickham2012-yr] transform the time coordinates into the space coordinates to plot the time series of different locations on the map. Calendar plots [@wang2020calendar] reconstructs time into calendar-based grid to discover weekday and weekend pattern. Projection, or linear combination, of variables summarises multivariate information into lower dimension to further digest. This section elaborates on the glyph map.

In \proglang{R}, \pkg{GGally} implements glyph maps through the \code{glyphs()} function. The function constructs a data frame with calculated position (\code{gx}. \code{gy}. \code{gid}) of each point on the time series using linear algebra (Equation 1 and 2 in @Wickham2012-yr). The data can then be piped into \code{ggplot} to create the glyph map as:

### 2022-02-08 main.tex
add transformation of calendar plot and projection

\hypertarget{sf}{%
\subsubsection{sf}\label{sf}}

\textcolor{red}{add content}

@@

\textcolor{red}{Glyph map + transforming on time (calendar plot) + projections (linear combination).}

Glyph map \citep{Wickham2012-yr} plots the time series as single glyph
on the map and in \proglang{R}, \pkg{GGally} implements the glyph map
through the \code{glyphs()} function. The function constructs a data
frame with calculated position (\code{gx}. \code{gy}. \code{gid}) of
each point on the time series using linear algebra (Equation 1 and 2 in
\citet{Wickham2012-yr}). The data can then be piped into \code{ggplot}
to create the glyph map as:

>>>>

\hypertarget{sf-and-s2}{%
\subsubsection{sf and s2}\label{sf-and-s2}}

Cubble is compatible with simple feature geometry list-column
(\code{sfc}) in \pkg{sf} and \code{s2_lnglat} vector in \pkg{S2} in the
nested form. After creating a cubble, analysts can add the geometry
vectors as needed to perform spatial operations:

\begin{CodeChunk}
\begin{CodeInput}
R> library(s2)
R> # create a cubble
R> cb <- climate_flat %>% 
+   cubble::as_cubble(key = id, index = date, coords = c(long, lat))
R> 
R> # western australia map
R> wa <- ozmaps::abs_ste %>% filter(NAME == "Western Australia")
R> 
R> # create a `s2_lnglat` vector based on long and lat and filter stations in wa
R> cb %>%
+   mutate(ll = s2_lnglat(long, lat)) %>% 
+   filter(s2_within(ll, wa))
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [3]: nested form
# bbox:     [115.97, -32.94, 117.18, -31.89]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name          wmo_id ts                 ll      
  <chr>       <dbl> <dbl> <dbl> <chr>          <dbl> <list>             <s2_lng>
1 ASN00009021 -31.9  116.  15.4 perth airport  94610 <tibble [366 x 4]> (115.97~
2 ASN00010311 -31.9  117. 179   york           94623 <tibble [366 x 4]> (116.76~
3 ASN00010614 -32.9  117. 338   narrogin       94627 <tibble [366 x 4]> (117.17~
\end{CodeOutput}
\end{CodeChunk}

\begin{CodeChunk}
\begin{CodeInput}
R> library(sf)
R> # simplify australia map
R> aus <- ozmaps::abs_ste %>% rmapshaper::ms_simplify(keep = 0.01)
R> 
R> dt <- cb %>%
+   mutate(
+     # create `sfc` column based on long and lat
+     ll = st_sfc(
+       purrr::map2(long, lat, ~st_point(c(.x, .y))),
+       crs = st_crs(aus)),
+     # append state multi-polygon based on the `sfc` created
+     state = aus$geometry[st_within(ll, aus, sparse = FALSE)],
+     # adopt a different projection: lambert conformal conic (EPSG:3112)
+     state = st_transform(state, crs = "EPSG:3112"))
R> 
R> dt
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name         wmo_id ts                        ll
  <chr>       <dbl> <dbl> <dbl> <chr>         <dbl> <list>           <POINT [°]>
1 ASN00009021 -31.9  116.  15.4 perth airpo~  94610 <tibble~ (115.9764 -31.9275)
2 ASN00010311 -31.9  117. 179   york          94623 <tibble~  (116.765 -31.8997)
3 ASN00010614 -32.9  117. 338   narrogin      94627 <tibble~ (117.1797 -32.9342)
4 ASN00014015 -12.4  131.  30.4 darwin airp~  94120 <tibble~ (130.8925 -12.4239)
5 ASN00015131 -17.6  134. 220   elliott       94236 <tibble~ (133.5407 -17.5521)
# ... with 1 more variable: state <MULTIPOLYGON [m]>
\end{CodeOutput}
\end{CodeChunk}

@@

Spatio-temporal transformation is a useful technique to extract
information form spatio-temporal data. Glyph maps \citep{Wickham2012-yr}
transform the time coordinates into the space coordinates to plot the
time series of different locations on the map. Calendar plots
\citep{wang2020calendar} reconstructs time into calendar-based grid to
discover weekday and weekend pattern. Projection, or linear combination,
of variables summarises multivariate information into lower dimension to
further digest. This section elaborates on the glyph map.

In \proglang{R}, \pkg{GGally} implements glyph maps through the
\code{glyphs()} function. The function constructs a data frame with
calculated position (\code{gx}. \code{gy}. \code{gid}) of each point on
the time series using linear algebra (Equation 1 and 2 in
\citet{Wickham2012-yr}). The data can then be piped into \code{ggplot}
to create the glyph map as:
